---
title: |
  | Automated QC of DWI 
  | with `eddy quad` and `MRTrix` metrics
  | Step 1 of 2
output:
  html_document:
    code_folding: hide

---

code written: 2020-07-12     
last ran: `r Sys.Date()`   

-----

```{r setup, include=FALSE}

#clean workspace
rm(list=ls())

#load ibraries
libraries <- c('tidyverse', 'reshape2', 'ggplot2', 'ggfortify') 
lapply(libraries, require, character.only = T) ; rm(libraries)

#bring in data 
mrtrix <- read.csv('../data/input/mrtrixResidualNoise.txt', header=F, sep='', col.names = c('residual', 'record_id')) 
eddy <- read.csv('../data/input/eddyMetrics.tsv', header=T, sep='\t') 

```

```{r data munging, warning=FALSE}

#merge the variables we are interested in
df <- merge(eddy[, c('subject_id', 'qc_mot_abs', 'qc_mot_rel', 'qc_outliers_pe', 'avg_snr_0', 'avg_cnr_1000')], mrtrix, 
            by.x='subject_id', by.y='record_id', all=TRUE)

#make sure all values are numeric
df[2:ncol(df)] <- sapply(df[2:ncol(df)],as.numeric)

#if blanks, turn to NA
df[df== ''] <- NA

```

__Correlations__. For data exploration purposes, we can investigate correlations between the 6 QC variables. Interestingly, in our past analyses, `Residual noise` was found to be uncorrelated with the eddy metrics; it seems to be contributing something unique. In our past analyses, all other correlations were as would be expected (e.g., high correlation between motion metrics).

```{r quality control}

#grab 6 variables of interest
pca_vars <- df[, c(
  "qc_mot_abs",         
  "qc_mot_rel" ,         
  "qc_outliers_pe" ,     
  "avg_snr_0",           
  "avg_cnr_1000",
  "residual")]

#rename the variables for the plot
names(pca_vars) <-  c(
  "Absolute motion", 
  "Relative motion",
  "Percent outliers",
  "Average SNR",
  "Average CNR", 
  "Residual noise")

#calculate correlation matrix
corr <- round(cor(pca_vars, use = "pairwise.complete.obs"), 2)

#melt the matrix
corr <- melt(corr)

#make plot
ggplot(corr, aes(Var1, Var2, fill=value)) +
  geom_tile() + 
  theme(legend.position = 'none',
        axis.title=element_blank(),
        panel.background = element_blank()) +
  geom_text(aes(Var1, Var2, label = value), color = "white", size = 4)

```

-----

__PCA.__ We performed a PCA, to determine how many PCs are required to describe a substantial portion of the variance. In our applications to SPINS, HCP, and POND data, we found that the first PC is the `eddy` metrics, and the second is `MRTrix`. We decided to include the first PC value across all 6 DWI QC metrics as a regressor for subsequent analyses.

```{r warning=FALSE}

#perform PCA 
(pca <- prcomp(pca_vars, center = TRUE, scale. = TRUE))

#review importance of components
summary(pca)

#visualize in plot
pcaPlot_fn <- function(pca_vars){
autoplot(prcomp(pca_vars, center = TRUE, scale. = TRUE),
         df, size=4.5, alpha=.6,
         loadings = TRUE,
         loadings.label = TRUE, loadings.label.size = 4, 
         loadings.colour='black', loadings.label.colour='black') +
  theme(legend.position='none',
        panel.background = element_blank()) +
  xlim(-.3,.3)
}

#run function 
pcaPlot_fn(pca_vars)

```

```{r}

#get the PC scores for all participants, into our df
cols <- rapply(pca, classes = 'matrix', how = 'list', f = function(x) x[, 1, drop = FALSE]) 
pca_col <- as.vector(cols$x)
df$qc_PC1 <- pca_col

#write out data
write.csv(df, paste0('../data/output/df_dwiQC_PC_', Sys.Date(), '.csv', sep=''), row.names = F) 

```

