---
title: |
  | Automated QC of SPINS DWI 
  | with `eddy quad` and `MRTrix`
output:
  html_document:
    code_folding: hide

---

code written: 2020-07-12     
last ran: `r Sys.Date()`   

-----

```{r setup, include=FALSE}

#clean environment
rm(list = ls(all.names = TRUE))

#load libraries
libraries <- c('ggplot2', 'reshape2', 'tidyverse', 'ggrepel', 'kableExtra') 
lapply(libraries, require, character.only = T) 

#bring in data
df <- read.csv(dir('../csvs/modified', full.names=T, pattern="^df_dwiQC")) #422

```

<br>
__Description.__ This script summarizes key QC metrics from the automated `eddy QC` tools, i.e., the eddy quad PDFs (Bastiani et al., _NeuroImage_, 2019), and `MRTrix` for the n=`r nrow(df)` participants who are eligible for inclusion.

<br>
__Thresholds for key eddy QC metrics.__ Our DWI working group suggested thresholds for key QC metrics, on the basis of analysis of subsets of the SPINS, POND, and HCP datasets. These thresholds were found to reliably predict a number of different visual QC ratings from different raters. The suggested thresholds were as follows:

```{r thresholds, echo=FALSE, message=FALSE}

#make a vector with metric names
metric_names <- c("Absolution motion", 
             "Relative motion", 
             "Percent outliers", 
             "Average SNR", 
             "Average CNR",
             "Residual noise")

#make a vector with threshold values set by John
thresholdFail <- c('>= 2 (mm)', 
                   '>= .5 (mm)', 
                   '>= 2 (%)',
                   '<= 20', 
                   '<= 1.5',
                   '>= 6')

#make a dataframe
threshold <- as.data.frame(rbind(metric_names, thresholdFail), row.names = F, header = F)
row.names(threshold) <- c('Metric', 'Suggested threshold')
colnames(threshold) <- NULL

#output in table
threshold %>%
  kable(align='c') %>%
  kable_styling()

```

<br>
__Visualizations of SPINS data quality.__ The distributions of the key QC metrics are visualized in the tabs below. The suggested thresholds are marked by the dashed vertical line. Participants with values exceeding the suggested thresholds are indicated by participant ID. 

```{r review distributions, warning=FALSE, fig.width=9.5, fig.height=4, message=FALSE}

#write a function for plotting automated QC metrics
qcPlot_fn <- function(yvar, threshold, operator){
  
#fancy fanagling to use variable and operator in ggplot
yvar <- eval(substitute(yvar), df)
operator_fn <- match.fun(operator)

#feed data into ggplot 
df %>% ggplot(aes(x=factor(0), y = get(yvar))) +
  
  #add jittered points 
  geom_jitter(width=.5, alpha=.4, size=5, pch=21) +

  #overlay violin plot
  geom_violin(trim = FALSE, fill='lightgrey', color='black', alpha=.2, color = NA) +

  #specify plot appearance
  coord_flip() +
  theme_classic() +
  ylab('') +
  xlab('') +
  theme(legend.position='none',
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.line=element_blank()) +

  #add cut-off score line
  geom_hline(yintercept = threshold, linetype = "dashed", color = "black") +
  
  #add a label to outliers
  geom_label_repel(data = subset(df, operator_fn(get(yvar), threshold)),
      aes(label = record_id), segment.colour = 'black', nudge_x =.2)

}

```
## {.tabset}

###Absolute motion 
```{r echo=FALSE, warning=FALSE, fig.width=9.5, message=FALSE}
qcPlot_fn('qc_mot_abs', 2, '>=')
```

###Relative motion
```{r echo=FALSE, warning=FALSE, fig.width=9.5, message=FALSE}
qcPlot_fn('qc_mot_rel', .5, '>=')
```

###Percent outliers
```{r echo=FALSE, warning=FALSE, fig.width=9.5, message=FALSE}
qcPlot_fn('qc_outliers_pe', 2, '>=')
```

###Average SNR
```{r echo=FALSE, warning=FALSE, fig.width=9.5, message=FALSE}
qcPlot_fn('qc_avg_snr',  20, '<=')
```

###Average CNR
```{r echo=FALSE, warning=FALSE, fig.width=9.5, message=FALSE}
qcPlot_fn('qc_avg_cnr', 1.5, '<=')
```

###Residual noise 
```{r echo=FALSE, warning=FALSE, fig.width=9.5, message=FALSE}
qcPlot_fn('qc_residual', 6, '>=')
```
</div>

-----

```{r data munging to set thresholds echo=FALSE, message=FALSE, results='hide'}

#make columns to indicate PASS/FAIL on basis of working group recommendations
 df$qc_eval_mot_abs <- ifelse(df$qc_mot_abs >= 2, 1, 0)
 df$qc_eval_mot_rel <- ifelse(df$qc_mot_rel >= .5,  1, 0)
 df$qc_eval_outliers_pe <- ifelse(df$qc_outliers_pe >= 2,  1, 0)
 df$qc_eval_avg_snr <- ifelse(df$qc_avg_snr <= 20,  1, 0)
 df$qc_eval_avg_cnr <- ifelse(df$qc_avg_cnr <= 1.5,  1, 0)
 df$qc_eval_residual <- ifelse(df$qc_residual >= 6,  1, 0)

#make global FAIL count
df$qc_score <- rowSums(df[, grep('eval', colnames(df))])

#identify participants with score of 2 or more -- these participants will be removed
exclude_qc <- paste('DWI_quality', df$record_id[which(df$qc_score >= 2)])

#write out names for removal from future analyses
write(exclude_qc, file = "..data/output/excludeParticipants.txt",
      ncolumns = 1,
      append = TRUE, sep = " ")

```

<br>
__Potential participant exclusion.__ The DWI working group recommended that participants exceeding threshold on 2 or more QC metrics should be removed from subsequent analyses. (Participants not exceeding this threshold will still undergo visual/non-automatic quality control and possibility be excluded for similar or unrelated reasons in that process.) If this decision were applied here, we would omit n=`r length(exclude_qc)` participants analysis, as follows:

```{r echo=FALSE}

#subset df to those who have a score of 2 or greater
df <- df[df$qc_score >= 2,]

#round values outside of the table
df <- df %>% mutate_if(is.numeric, round, digits=2)

#make a summary table
df %>%
  mutate(
    qc_mot_abs = cell_spec(qc_mot_abs, "html", color='white', background=ifelse(qc_mot_abs >=.2, "red", "green")),
    qc_mot_rel = cell_spec(qc_mot_rel, "html", color='white', background=ifelse(qc_mot_rel >= .5, "red", "green")),
    qc_outliers_pe = cell_spec(qc_outliers_pe, "html", color='white', background=ifelse(qc_outliers_pe >=2, "red", "green")),
    qc_avg_snr = cell_spec(qc_avg_snr, "html", color='white', background=ifelse(qc_avg_snr <=20, "red", "green")),
    qc_avg_cnr = cell_spec(qc_avg_cnr, "html", color='white', background=ifelse(qc_avg_cnr <=1.5, "red", "green")),
    qc_residual = cell_spec(qc_residual, "html", color='white', background=ifelse(qc_residual >=6, "red", "green"))) %>%
  select(record_id,
         qc_mot_abs, qc_mot_rel, qc_outliers_pe,
         qc_avg_snr, qc_avg_cnr, qc_residual, qc_score) %>%
  kable(format='html', escape=F, align='c', 
        col.names = c('Participant', 'Absolute motion', 'Relative motion', 'Percent outliers', 'Average SNR', 'Average CNR', 'Residual noise', 'Flag count')) %>%
  kable_styling('striped')

```

